Climatic warming strengthens a positive feedback\\ between alpine shrubs and fire \\[1cm] \huge Supplementary Material

### 2017-01-03 manuscript/supplementary.Rnw
Problem with compiling pdf with line numbers using older versions of TeX Live (see #3 and annakrystalli/OpenConBerlin_ReproHack#2). Temp solution is to remove line numbers
Final edits for GCB

\documentclass[9pt,onecolumn,twoside, lineno]{gcb}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment. 
\pdfminorversion=4
\templatetype{gcbresearcharticle} % Choose template 
\usepackage{float}

@@

\begin{document}
\suppmaketitle
\tableofcontents
\listoffigures

\section{Study site}
Our research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia. The undulating terrain ranges in altitude from 1400 to 2000 m a.s.l and contains a well-developed mantle of soil. Mean annual temperature is approximately 6\textdegree C, with persistent snow cover from June to September. Annual precipitation is 1200 mm, yet droughts occur at roughly decadal intervals \citep{Wahren:2013ji}. The vegetation is a mosaic of Eucalypt-woodland and treeless vegetation, with the latter largely consisting of heathlands and tussock grasslands, but also herbfields and wetlands. The treeline is irregular (due to cold air drainage), occurring at altitudes ranging from 1600 to 1700 m a.s.l. Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003 \citep{Williams:2006tl,Camac:2013eb}. Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests \citep{Williams:2008ba}.

\section{Open Top Chamber Experiment}
We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by tussock grass and shrub seedling recruitment, growth and mortality; 2) how proximity to grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.

@@

Four seedlings of each species were then transplanted into 70 cm$^2$ of burnt bare ground in the center of each experimental plot. An additional four \textit{Grevillea} seedlings were transplanted into occuring post-fire inter-tussock gaps of varying size surrounding this bare ground. Only \textit{Grevillea} seedlings were transplanted into inter-tussock gaps because of space constraints but also because it is the dominant open heathland shrub species (i.e. occurs closest to the shrub/grassland boundary) and is therefore the most likely to invade grasslands. The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig. \ref{figs1}). However, the two sets of seedlings (i.e. bare ground vs inter-tussock) were treated separately because they were spatially confounded (e.g. center vs edge of the 1 m$^2$ opening) and addressed different questions. Seedlings in bare ground patches (i.e. the center of the plot) were used to assess whether warmer post-fire conditions altered growth and mortality rates in four dominant alpine shrub species. By comparison, \textit{Grevillea} seedlings growing in inter-tussock gaps were used to determine whether inter-tussock gap size altered growth and mortality rates, and whether such effects were modified by warmer conditions.

To simulate near-term warmer conditions indicated by the IPCC \citep{IPCC:2013vv}, we used Open Top Chambers (OTCs) \citep{Molau:1996vc}. OTCs were hexagonal, 58 cm tall with an open-top diameter of 110 cm and base diameter of 168 cm. The chambers were constructed following the International Tundra Experiment (ITEX) protocols \citep{Molau:1996vc} using sun-lite: a lightweight shatter-resistant fiberglass with high light transmittance (86\%) and low infrared transmittance (\textless 5\%; Solar Components Corporation, USA).

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) installed in four randomly chosen control plots and four OTC plots. Each Micro Station measured ambient temperature (10 cm above ground), relative humidity (10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions \citep{IPCC:2013vv} (Supplementary Fig. \ref{figs6}-\ref{figs8}). OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C and \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean','grand_mean')}\textdegree C, respectively. Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C). Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

@@

\subsection{Data inputs for growth, mortality and recruitment models}
In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$. However, because some distances were censored (i.e. $\geq$ 26 cm), these data were incomplete. We estimated these censored data using a censored model (see Modelling censored Poa Distances). Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations \citep{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.

@@

The study used 40 open heathland sites \citep{Camac:2013eb} established immediately after the 2003 fires \citep{Williams:2006tl}. All sites have been free of grazing by cattle since 2003. Thirty sites were burnt, while the remaining ten were unburnt, likely for more than 70 years \citep{Camac:2013eb}. This disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires. These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitude from \Sexpr{round(min(complete_field_survey_data$altitude_m))} to \Sexpr{round(max(complete_field_survey_data$altitude_m))} m a.s.l. Sites were at least 0.25 ha and more than 500 m apart. In the summer of 2011-12, 9-years post-fire, seedling density in each site was estimated using forty 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects. Transects were positioned in the center of each site and were parallel to each other with exactly 10 m between transect lines. This effectively created a grid of 40 plots (10 plots~\texttimes~4 transects) per site, subsampling an area of 2000 m$^2$.

@@

\section{Data analysis}
In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset \citep{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")} \citep{stan:2015xx}. For each model, three independent chains were run. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic \citep{Gelman:1992ts, Brooks:1998ju}. In all cases modelled parameters converged within 1000 iterations. After discarding the first 1000 iterations as ‘warm up’, a further 1000 samples were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. To ensure all covariate effects were identifiable we removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.

@@

\bibliography{refs.bib}
\pagebreak

\section{Supplementary Figures}

\begin{figure}[h]
\centering
\includegraphics{figures/figS1.pdf}
\caption{Diagram outlining positions of transplanted seedling in an OTC plot. The inner white square represents the burnt bare ground patch used to assess heights of four dominant alpine shrub species. The grey area contains \textit{Grevillea australis} seedlings randomly transplanted into burnt inter-tussock gaps. The area beyond the grey section was subject to OTC panel sheltering (i.e. it was not within the 110 cm opening of the chamber) and therefore was not used in the experiment.} 
\label{figs1}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS2.pdf}
\caption{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (twig diameter), altitude and Topographic Wetness Index (TWI). (a) Centered and standardized model coefficients (on the log scale); and effects of (b) burning, (c) fire severity, (d) altitude and (e) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs2}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS3.pdf}
\caption{Average number of \textit{Grevillea australis} recruits in warmed (red) and control (blue) conditions in 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}
\label{figs3}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS4.pdf}
\caption{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs. Red and Blue lines represent mean ($\pm$ 95\% CI) observed heights at each May census in warmed and control plots, respectively. Orange and light blue lines represent individual growth trajectories in warmed and unwarmed plots, respectively.} 
\label{figs4}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS5.pdf}
\caption{Mean maximum seedling height along fire severity (twig diameter), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires. Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses. (a \& b) Centered and standardized model coefficients (on the log scale); and effects of (c \& d) fire severity, (e \& f) altitude and (g \& h) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs5}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS6.pdf}
\caption{Mean daily microclimate difference (OTC $-$ Control). (a) Ambient temperature, (b) soil temperature, (c) volumetric water content and (d) relative humidity over experimental period. Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
\label{figs6}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS7.pdf}
\caption{Mean minimum daily microclimate difference (OTC $-$ Control). (a) Ambient temperature, (b) soil temperature, (c) volumetric water content and (d) relative humidity over experimental period. Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
\label{figs7}
\end{figure}
\clearpage

\begin{figure}[!hb]
\centering
\includegraphics{figures/figS8.pdf}
\caption{Mean maximum daily microclimate difference (OTC $-$ Control). Ambient temperature (a), soil temperature (b), volumetric water content (c) and relative humidity over experimental period (d). Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: amplitude declines in later years are due to shading effects caused by increased shrub growth.}
\label{figs8}
\end{figure}

>>>>

\documentclass[9pt,onecolumn,twoside]{gcb}
% Use the lineno option to display guide line numbers if required.
% Note that the use of elements such as single-column equations
% may affect the guide line number alignment. 
\pdfminorversion=4
\templatetype{gcbresearcharticle} % Choose template 
\usepackage{float}

@@

\begin{document}
\suppmaketitle
\tableofcontents
\clearpage
\listoffigures
\clearpage
\section{Study site}
Our research sites were on the Bogong High Plains, a series of alpine and high subalpine plateaux in the Alpine National Park, approximately 250 km northeast of Melbourne, Australia. The undulating terrain ranges in altitude from 1400 to 2000 m a.s.l and contains a well-developed mantle of soil. The most widespread soil type is ``alpine humus soil'', a friable, organic loam up to 1 m deep \citep{Williams:2014ud}. Mean annual temperature is approximately 6\textdegree C, with persistent snow cover from June to September. Annual precipitation is 1200 mm, yet droughts occur at roughly decadal intervals \citep{Wahren:2013ji}. The vegetation is a mosaic of Eucalypt-woodland and treeless vegetation, with the latter largely consisting of heathlands and tussock grasslands, but also herbfields and wetlands. The treeline is irregular (due to cold air drainage), occurring at altitudes ranging from 1600 to 1700 m a.s.l. Approximately 50\% of the alpine ‘treeless’ landscape consists of heathlands, most of which were extensively burned by a 400,000 ha wildfire in 2003 \citep[Fig. \ref{figS1};][]{Williams:2006tl,Camac:2013eb}. Large wildfires in Australian alpine landscapes have occurred once or twice per century and typically occur during periods of extended regional drought coupled with severe fire weather and multiple ignitions in the surrounding montane forests \citep{Williams:2008ba}.

\section{Open Top Chamber Experiment}
We used a warming experiment to examine: 1) how warmer post-fire conditions influence rates of post-fire gap-infilling by tussock grass as well as shrub seedling recruitment, growth and mortality; 2) how proximity to tussock grass affects shrub seedling demographic rates; and 3) whether such effects change under warmer conditions post-fire conditions.

@@

Four seedlings of each species were then transplanted into 70 cm$^2$ of burnt bare ground in the center of each experimental plot. An additional four \textit{Grevillea} seedlings were transplanted into occuring post-fire inter-tussock gaps of varying size surrounding this bare ground. Only \textit{Grevillea} seedlings were transplanted into inter-tussock gaps because of space constraints but also because it is the dominant open heathland shrub species (i.e. occurs closest to the shrub/grassland boundary) and is therefore the most likely to invade grasslands. The combination of bare ground and inter-tussock seedlings meant that each experimental plot had a total of 20 seedlings per 1 m$^2$ (Fig. \ref{figs5}). However, the two sets of seedlings (i.e. bare ground vs inter-tussock) were treated separately because they were spatially confounded (e.g. center vs edge of the 1 m$^2$ opening) and addressed different questions. Seedlings in bare ground patches (i.e. the center of the plot) were used to assess whether warmer post-fire conditions altered growth and mortality rates in four dominant alpine shrub species. By comparison, \textit{Grevillea} seedlings growing in inter-tussock gaps were used to determine whether inter-tussock gap size altered growth and mortality rates, and whether such effects were modified by warmer conditions.

To simulate near-term warmer conditions indicated by the IPCC \citep{IPCC:2013vv}, we used Open Top Chambers (OTCs; Fig. \ref{figS2}) \citep{Molau:1996vc}. OTCs were hexagonal, 58 cm tall with an open-top diameter of 110 cm and base diameter of 168 cm. The chambers were constructed following the International Tundra Experiment (ITEX) protocols \citep{Molau:1996vc} using sun-lite: a lightweight shatter-resistant fiberglass with high light transmittance (86\%) and low infrared transmittance (\textless 5\%; Solar Components Corporation, USA).

We measured changes in microclimate using Onset Micro Stations (Onset Computer Corporation, Bourne, MA, USA) installed in four randomly chosen control plots and four OTC plots. Each Micro Station measured ambient temperature (10 cm above ground), relative humidity (10 cm above ground), soil temperature (3 cm below ground) and volumetric soil moisture (between 3 and 10 cm) at hourly intervals for the entire duration of the experiment. Across five growing seasons, OTCs simulated warmer conditions at the lower end of IPCC predictions \citep{IPCC:2013vv} (Supplementary Fig. \ref{figs6}-\ref{figs8}). OTCs passively increased average daily ambient and soil temperatures by \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean','grand_mean')}\textdegree C. Minimum and maximum temperatures were also raised in both ambient air (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CAmbient_Temp','mean_max','grand_mean')}\textdegree C) and soil (min: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_min','grand_mean')}\textdegree C; max: \Sexpr{micro_clim(overall_microclimate_summary, 'CTemp_3cmBG','mean_max','grand_mean')}\textdegree C). Chambers only marginally decreased soil moisture by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CMoisture3to10cmBG','mean','grand_mean'))}\% and relative humidity by \Sexpr{abs(micro_clim(overall_microclimate_summary, 'CRH','mean','grand_mean'))}\%. Because OTCs strongly increase temperatures relative to other measures, throughout we refer to it as a ‘warming treatment’.

@@

\subsection{Data inputs for growth, mortality and recruitment models}
In each model, OTC treatment, $OTC$, was a binary (0 = Control, 1 = OTC) plot level covariate. By contrast, inter-tussock gap radius, $Gap_{i,t}$, was a continuous covariate on the log-scale measured for each individual seedling in each census $t$. However, because some distances were censored (i.e. $\geq$ 26 cm), these data were incomplete. We estimated these censored data using a censored model (see: Modelling censored Poa distances). Once true distances were estimated for censored data, we estimated the median inter-tussock gap radius by log-transforming the data and then calculating the median cardinal distance to nearest tussock, in each census, for each inter-tussock seedling. Median inter-tussock gap radius ranged from \Sexpr{range_init_gap(tussock_otc_growth_data)[1]} cm to \Sexpr{round(range_init_gap(tussock_otc_growth_data)[2])} cm in both control and OTC plots. In order to compare the magnitude of effects between OTC treatment and inter-tussock gap radius we centered it on zero and divided by two standard deviations \citep{Gelman:2007te}. The advantage of centering and standardizing is that it allows for easier interpretation, with intercepts interpreted as responses to average conditions and slope terms as partial dependencies conditional on other continuous variables being at their mean.

@@

The study used 40 open heathland sites \citep[Fig. \ref{figS9};][]{Camac:2013eb} established immediately after the 2003 fires \citep{Williams:2006tl}. All sites have been free of grazing by cattle since 2003. Thirty sites were burnt, while the remaining ten were unburnt, likely for more than 70 years \citep{Camac:2013eb}. This disparity in sampling effort between burnt and unburnt was due to very few open heathland remaining unburnt after both the 2003 and 2007 fires. These sites encompassed most of the Bogong High Plains (120 km$^2$) and ranged in altitude from \Sexpr{round(min(complete_field_survey_data$altitude_m))} to \Sexpr{round(max(complete_field_survey_data$altitude_m))} m a.s.l. Sites were at least 0.25 ha and more than 500 m apart. In the summer of 2011-12, 9-years post-fire, seedling density in each site was estimated using forty 1 m$^2$ quadrats, evenly distributed in groups of 10 along four 50 m transects. Transects were positioned in the center of each site and were parallel to each other with exactly 10 m between transect lines. This effectively created a grid of 40 plots (10 plots by 4 transects) per site, subsampling an area of 2000 m$^2$.

@@

\section{Data analysis}
In total we built 11 hierarchical models to examine factors influencing post-fire recovery of tussock grass and shrub seedling establishment, growth and mortality. These models are well suited to datasets that have a hierarchical structure (e.g. plots within sites) because they can account for observation error and partition both explained and unexplained variation at multiple levels of a dataset \citep{Gelman:2007te, Kery:2012wl}. For each model we used Bayesian inference and fitted models in R \Sexpr{paste(version$major, version$minor, sep='.')} using package \texttt{rstan} \Sexpr{packageVersion("rstan")} \citep{stan:2016xx}. For each model, three independent chains were run. Convergence was assessed through both visual inspection of chains and reference to the Brooks-Gelman-Rubin convergence diagnostic \citep{Gelman:1992ts, Brooks:1998ju}. In all cases modelled parameters converged within 1000 iterations. After discarding the first 1000 iterations as ‘warm up’, a further 1000 samples were taken from the joint posterior. Below we describe the model structure, parameters and priors used in these models. To ensure all covariate effects were identifiable we removed each covariate and visually inspected whether the direction and magnitude of effects were severely altered; none were. Data and source code for reproducing analysis and figures is available at: \url{https://github.com/jscamac/Alpine_Shrub_Experiment}.

@@

\bibliography{refs.bib}

\clearpage
\section{Supplementary Figures}

\begin{figure}[!h]
\centering
\includegraphics{figures/figS1.jpg}
\caption{Approximately 400,000 ha of the Victorian Alpine National Park was burned by the wildfires in 2003.} 
\label{figs1}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS2.jpg}
\caption{Open-Top-Chambers in \textit{Grevillea australis} and \textit{Poa hiemata} dominated open heathland.} 
\label{figs2}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS3.jpg}
\caption{Shrubs are the most flammable component in Australian alpine 'treeless' vegetation \citep{Williams:2006tl}. Burned patches were shrubs surrounded by lightly scorched tussock grass.} 
\label{figs3}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS4.jpg}
\caption{Setting up of experimental plots. (a): unburned \textit{Grevillea australis}; (b): \textit{Grevillea} after being burned; (c) \textit{Grevillea} remains removed leaving burned bare ground patch. Transplant holes were then created using hamilton planter (note: photo does not show holes for inter-tussock seedlings; (d) all seedlings planted within 1 m$^2$.} 
\label{figs4}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS5.pdf}
\caption{Diagram outlining positions of transplanted seedling in an OTC plot. The inner white square represents the burnt bare ground patch used to assess heights of four dominant alpine shrub species. The grey area contains \textit{Grevillea australis} seedlings randomly transplanted into burnt inter-tussock gaps. The area beyond the grey section was subject to OTC panel sheltering (i.e. it was not within the 110 cm opening of the chamber) and therefore was not used in the experiment.} 
\label{figs5}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS6.pdf}
\caption{Mean daily microclimate difference (OTC $-$ Control). (a) Ambient temperature, (b) soil temperature, (c) volumetric water content and (d) relative humidity over experimental period. Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
\label{figs6}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS7.pdf}
\caption{Mean minimum daily microclimate difference (OTC $-$ Control). (a) Ambient temperature, (b) soil temperature, (c) volumetric water content and (d) relative humidity over experimental period. Red lines = mean difference across all growing seasons (when chambers were applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: declines in amplitude in later years are due to shading effects caused by increased shrub size.}
\label{figs7}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS8.pdf}
\caption{Mean maximum daily microclimate difference (OTC $-$ Control). Ambient temperature (a), soil temperature (b), volumetric water content (c) and relative humidity over experimental period (d). Red lines = mean difference across all growing seasons (when chambers where applied); blue = no difference between OTCs and controls. Shaded areas are the winter snow periods when OTCs were not over plots. Note: amplitude declines in later years are due to shading effects caused by increased shrub growth.}
\label{figs8}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS9.png}
\caption{Distribution of open heathland sites across the Bogong high plains. Red points: sites burned by the 2003 wildfires; Blue points: unburned sites. White areas signify alpine treeless vegetation.}
\label{figs9}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS10.jpg}
\caption{Example of a shrub severely burned by the 2003 wildfires.}
\label{figs10}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS11.pdf}
\caption{Mean \textit{Asterolasia trymalioides} seedling density along gradients of burning, fire severity (MTD), altitude and Topographic Wetness Index (TWI). (a) Centered and standardized model coefficients (on the log scale); and effects of (b) burning, (c) fire severity; as measured by minimum twig diameter (MTD), (d) altitude and (e) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs11}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS12.pdf}
\caption{Average number of \textit{Grevillea australis} recruits in warmed (red closed circle) and control (blue open circle) conditions in 2011 and 2012. Errors bars are 95\% Bayesian Credible Intervals.}
\label{figs12}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS13.pdf}
\caption{Observed seedling diameter growth trajectories of four dominant Australian alpine shrubs. Thick lines represent mean ($\pm$ 95\% CI) observed diameters in warmed (red solid line/closed circles) and control (blue dashed line/open circles) at each May census in warmed and control plots, respectively. Thin lines represent individual growth trajectories in warmed (orange) and control (light blue) conditions.} 
\label{figs13}
\end{figure}
\clearpage

\begin{figure}[!h]
\centering
\includegraphics{figures/figS14.pdf}
\caption{2012 mean maximum seedling height along fire severity (MTD), altitude and Topographic Wetness Index (TWI) gradients in areas burnt by the 2003 wildfires. Left panels = \textit{Asterolasia trymalioides} responses, Right panels = \textit{Grevillea australis} responses. (a \& b) Centered and standardized model coefficients (on the log scale); and effects of (c \& d) fire severity; as measured by minimum twig diameter (MTD), (e \& f) altitude and (g \& h) Topographic Wetness Index (TWI), in areas burnt by the 2003 fires. All error bars and shaded areas indicate 95\% Bayesian Credible Intervals.}
\label{figs14}
\end{figure}
\clearpage
